---
title: "Phineas and Ferb Data Visualizations"
author: "Phillip Sanderell"
date: "10/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages

```{r}
library(tidyverse)
library(showtext)
```

# Import data

```{r}
songs <- read_csv("pnf_songs.csv")
```

# Data visualizations

Now comes the fun part, we have data and now we can visualize it to try to make some summarizations or inferences.

## Setting some formating presets to variables

```{r}
small_plot_text <- element_text(family = "Graphite Std Bold", 
                                 size = 15,
                                 color = "black")

big_plot_text <- element_text(face = "bold", 
                                  family = "Ferbtastic",
                                  size = 30)
```


## What are the most common song genres?

```{r}
songs %>% 
  mutate(genre = str_to_lower(genre)) %>% 
  separate(genre, into = c("g1", "g2", "g3", "g4", "g5"),
           sep = "[,/]") %>% 
  pivot_longer(g1:g5, "col", values_to = "genre") %>% 
  mutate(genre = str_trim(genre)) %>% 
  filter(!is.na(genre)) %>% 
  mutate(season = substr(pc, 1, 1)) %>% 
  group_by(genre) %>% 
  mutate(genre_count = n()) %>% 
  filter(genre_count > 10) %>% 
  count(genre, season, sort = TRUE) %>% 
  na.omit(season) %>% 
  group_by(genre) %>% 
  mutate(genre_count = sum(n)) %>% 
  ungroup() %>% 
  mutate(genre = fct_reorder(genre, genre_count)) %>% 
  mutate(season = case_when(season == "M" ~ "movie", 
                            TRUE ~ season)) %>% 
  mutate(season = fct_relevel(season, "movie", "4", "3", "2", "1")) %>% 
  ggplot(aes(y = genre, x = n, fill = season)) + 
  geom_col(color = "black") + 
  theme_minimal() + 
  scale_fill_brewer(palette = "YlOrRd", 
                    guide = guide_legend(reverse = TRUE)) + 
  labs(title = "Pop, Rock, and Show Tunes are the most\ncommon genre of P&F songs",
       caption = "\nData from phineasandferb.fandom.com  |  Scraped and Visualized by @PhilliPhresh") + 
  theme(axis.title = element_blank(), 
        panel.grid.major = element_blank(), 
        plot.title.position = "plot",
        plot.background = element_rect(fill = "#ffeac9"),
        plot.title = big_plot_text,
        axis.text = small_plot_text,
        legend.text = small_plot_text,
        legend.title = small_plot_text, 
        plot.caption = element_text(hjust = 1),
        plot.caption.position = "plot", 
        legend.direction = "horizontal", 
        legend.position = c(.65, .15))

ggsave("prevelance of phineas and ferb genres.png", width = 8, height = 5)
```

## Does the length of songs differ by season?

```{r}
songs %>% 
  mutate(genre = str_to_lower(genre)) %>% 
  mutate(length = str_extract(length, "[:digit:]+[:punct:][:digit:]+")) %>% 
  mutate(length = paste0("00:0", length) %>% parse_time()) %>% 
  mutate(season = substr(pc, 1, 1)) %>% 
  mutate(season = case_when(season == "M" ~ "movies", 
                            TRUE ~ season)) %>% 
  filter(!is.na(season)) %>% 
  mutate(season = fct_relevel(season, "1", "2", "3", "4", "movies")) %>%
  ggplot(aes(season, length, fill = season)) + 
  geom_violin(draw_quantiles = c(.25, .5, .75), kernel = "triangular",
              alpha = .9) + 
  annotate("text", x = 5.75, y = 2*60 + 30, label = "75th percentile", 
           hjust = 0, family = "Graphite Std Bold") + 
  geom_curve(aes(x = 5.75, y = 2*60 + 30, xend = 5.35, yend = 2*60),
              arrow = arrow(length = unit(0.03, "npc")), curvature = .4) + 
  annotate("text", x = 5.75, y = 1.75*60, label = "50th", 
           hjust = 0, family = "Graphite Std Bold") + 
  geom_curve(aes(x = 5.75, y = 1.75*60-1, xend = 5.45, yend = 1.75*60-1),
              arrow = arrow(length = unit(0.03, "npc")), curvature = 0) +
  annotate("text", x = 5.75, y = 1*60, label = "25th", 
           hjust = 0, family = "Graphite Std Bold") + 
  geom_curve(aes(x = 5.75, y = 1*60, xend = 5.3, yend = 1.5*60 - 5),
              arrow = arrow(length = unit(0.03, "npc")), curvature = -.4) +
  labs(title = "Length of songs in P&F seem to get longer\nand more predictible over time",
       x = NULL,
       y = NULL, 
       caption = "\nData from phineasandferb.fandom.com  |  Scraped and Visualized by @PhilliPhresh") + 
  expand_limits(x = 7.5) +
  scale_y_continuous(breaks = c(0, 1*60, 2*60, 3*60),
                     labels = c("", "1 minute", "2 minutes", "3 minutes")) +
  scale_x_discrete(breaks = c("1", "2", "3", "4", "movies"),
                   labels = c("Season 1", "2", "3", "4", "Movies")) + 
  scale_fill_brewer(palette = "YlOrRd", 
                    guide =  NULL, 
                    direction = -1) + 
  theme_minimal() + 
  theme(plot.title.position = "plot", 
        plot.title = big_plot_text, 
        axis.text = small_plot_text, 
        axis.title = small_plot_text,
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.background = element_rect(fill = "#ffeac9"),
        panel.grid.major.y = element_line(color = "#e0c69d"))

ggsave("prevelance of phineas and ferb song length.png", 
       width = 8, height = 5)
```

Does the IMDb rating distribution differ by genre?

```{r}
songs %>% 
  mutate(genre = str_to_lower(genre)) %>% 
  separate(genre, into = c("g1", "g2", "g3", "g4", "g5"),
           sep = "[,/]") %>% 
  pivot_longer(g1:g5, "col", values_to = "genre") %>% 
  mutate(genre = str_trim(genre)) %>% 
  filter(!is.na(genre)) %>% 
  mutate(med_overall_rating = median(imdb_rating, na.rm = TRUE)) %>% 
  group_by(genre) %>% 
  summarise(med_overall_rating = med_overall_rating, 
            med_imdb_rating = median(imdb_rating, na.rm = TRUE),
            std_imdb_rating = var(imdb_rating, na.rm = TRUE)^.5, 
             n_genre = n(),
            imdb_rating_diff = med_imdb_rating - med_overall_rating) %>% 
  distinct() %>% 
  filter(n_genre > 10) %>% 
  mutate(abs_rating_diff = abs(imdb_rating_diff)) %>% 
  slice_max(abs_rating_diff, n = 10) %>% 
  ungroup() %>% 
  mutate(genre = fct_reorder(genre, n_genre * 
                               if_else(imdb_rating_diff<0, -1, 1))) %>% 
  mutate(genre = fct_reorder(genre, med_imdb_rating)) %>%
  ggplot(aes(x = med_imdb_rating, y = genre)) + 
  geom_segment(aes(x = med_overall_rating, xend = med_imdb_rating,
                   y = genre, yend = genre)) + 
  geom_point(aes(size = n_genre), fill = "#F3390E", shape = 21) + 
  
  geom_curve(xend = 8.15, yend = 4.5, x = 8.16, y = 3.5,
             arrow = arrow(length = unit(0.03, "npc")), curvature = -.4) + 
  annotate(geom = "label", x = 8.16 + .005, y = 3.5, 
           label = "Larger dots mean\nmore episodes with\nthat genre of song",
           hjust = 0, lineheight = .8, family = "Graphite Std Bold",
           fill = "#FFFEAE") + 
  
  labs(title = "Episodes with rock songs have higher ratings",
       subtitle = "Episodes with pop or holiday songs have lower ratings",
       y = NULL,
       x = "Median IMDb Rating\n(for episodes with songs in various genres)",
       caption = "\nData from phineasandferb.fandom.com  |  Scraped and Visualized by @PhilliPhresh") +
  guides(size = "none") + 
  theme_minimal() + 
  scale_x_continuous(breaks = c(8, 8.10, 8.20)) + 
  theme(plot.title.position = "plot",
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = big_plot_text,
        axis.text = small_plot_text,
        axis.title = small_plot_text,
        plot.subtitle = element_text(face = "bold", 
                                  family = "Ferbtastic",
                                  size = 17),
        plot.background = element_rect(fill = "#ffeac9"),
        panel.grid.major.y = element_line(color = "#e0c69d"))

ggsave("ratings of phineas and ferb episodes.png", 
       width = 8, height = 5)
```

Can we safely assume IMDb ratings are normally distributed?

```{r}
songs %>% 
  distinct(ep_title, .keep_all = TRUE) %>% 
  ggplot(aes(imdb_rating)) +
  geom_density()
```

I would say yes!

So now we can try a variation of the plot above only use a TIE Fighter plot instead of a lolipop plot.

```{r}
songs %>% 
  mutate(genre = str_to_lower(genre)) %>% 
  separate(genre, into = c("g1", "g2", "g3", "g4", "g5"),
           sep = "[,/]") %>% 
  pivot_longer(g1:g5, "col", values_to = "genre") %>% 
  mutate(genre = str_trim(genre)) %>% 
  filter(!is.na(genre)) %>% 
  filter(!is.na(imdb_rating)) %>% 
  # filter(genre == "pop") %>% #testing to see if group_by is working,
  # I don't think it is
  nest(data = -genre) %>% 
  mutate(ci = map(data, ~ confintr::ci_quantile(x = pull(., imdb_rating)) %>% .$interval)) %>% 
  unnest(data) %>% 
  unnest(ci) %>% 
  filter(ci != Inf & ci != -Inf) %>% 
  mutate(column = row_number() %% 2) %>% 
  pivot_wider(names_from = column, values_from = ci) %>% 
  rename(lb = `1`, ub = `0`) %>% 
  group_by(genre) %>% 
  summarise(med = median(imdb_rating, na.rm = TRUE),
            lb = median(lb, na.rm = TRUE),
            ub = median(ub, na.rm = TRUE),
            n = n()) %>% 
  mutate(color = if_else(genre %in% c("country", "rock", "hip-hop"),
                         "red", "orange")) %>% 
  filter(n > 7) %>% 
  mutate(genre = fct_reorder(genre, med)) %>% 
  ggplot(aes(med, genre, fill = color)) +
  geom_errorbar(aes(xmin = lb, xmax = ub)) + 
  geom_point(aes(size = n), shape = 21) + 
  
  annotate("text", label = "lower bound", x = 8, y = 14, 
           hjust = .5, family = "Graphite Std Bold") +
  annotate("text", label = "median", x = 8.25, y = 14, 
           hjust = .5, family = "Graphite Std Bold") +
  annotate("text", label = "upper bound", x = 8.7, y = 14, 
           hjust = .5, family = "Graphite Std Bold") +
  
  geom_curve(xend = 7.97, yend = 4, x = 7.9, y = 5, color = "black",
             arrow = arrow(length = unit(0.03, "npc")), curvature = .2) + 
  annotate(geom = "label", x = 7.6, y = 6.5, 
           label = "Larger dots mean\nmore episodes with\nthat genre of song",
           hjust = 0, lineheight = .8, family = "Graphite Std Bold",
           fill = "#FFFEAE") + 
  
  guides(size = "none", fill = "none") +
  theme_minimal() +
  scale_fill_identity() + 
  expand_limits(y = 15) +
  scale_x_continuous(limits = c(7.6, 9), 
                     breaks = c(seq(7.8, 8.6, .2), 8.84),
                     labels = c("7.8", "8.0", "8.2", "8.4", "8.6", "Stars on IMDb")) +
  theme(axis.title.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title.position = "plot",
        plot.title = big_plot_text,
        axis.text = small_plot_text,
        axis.title = small_plot_text,
        panel.grid.major.y = element_line(color = "#e0c69d"),
        plot.background = element_rect(fill = "#ffeac9")) +
  labs(title = "The only genres with significantly different\nratings were Rock, Country, and Hip-Hop",
       x = "(for episodes with songs of that genre)",
       caption = "\nData from phineasandferb.fandom.com  |  Scraped and Visualized by @PhilliPhresh")

ggsave("ratings of phineas and ferb episodes TIE fighter plot.png", 
       width = 8, height = 5)
```

```{r}
songs_I_like <- 
  c("When We Didn't Get Along",
  "City of Love",
  "Let's Take A Rocketship To Space",
  "E.V.I.L. B.O.Y.S.",
  "Busted",
  "Flawless Girl",
  "Disco Miniature Golfing Queen",
  "S'Winter",
  "Big Honkin' Hole in My Heart",
  "Ain't Got Rhythm",
  "You Snuck Your Way Right Into My Heart",
  "Little Brothers",
  "Chains on Me",
  "Yodel Odel Obey Me",
  "Izzy's Got the Frizzies",
  "There's a Platypus Controlling Me",
  "Dance Baby",
  "Aerial Area Rug",
  "Double Dutch Song",
  "Waggle Dance",
  "In the Empire",
  "When Tomorrow is This Morning Again",
  "O.W.C.A. Files",
  "My Undead Mummy")

songs %>% 
  arrange(air_date) %>% 
  mutate(songs_I_really_like = if_else(song_title %in% songs_I_like, "I really like it", "It's okay")) %>% 
  ggplot(aes(air_date, imdb_rating)) +
  geom_point() +
  gghighlight::gghighlight(songs_I_really_like == "I really like it")
```

