---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aren't you a little old to be watching Phineas and Ferb?

Yes, yes I am. I love Phineas and Ferb. From the jokes to the songs to the inators, everything about it brings me joy. Going through a recent watch on Disney+ inspired me to analyze Phineas and Ferb data, specifically, song data.

# Where did you get the data?

While making the plots (see below) was the more exciting part, gathering the data was definitely the more challenging and time-consuming part of this project. After looking around various websites, I ended up deciding to web scrape from the [Phineas and Ferb Fandom page](https://phineasandferb.fandom.com/wiki/Phineas_and_Ferb_Wiki). To pull IMDB ratings, I obviously also had to scrape data from [IMDb](https://www.imdb.com/title/tt0852863/episodes?season=1).

Since Fandom is a Wiki site, the webpages scraped were not always in a consistent format and some assumptions needed to be made. The three largest assumptions made were:

1.  The first episode to be listed in a song's narrative description is the episode it first aired in.
2.  All the data taken from the Fandom Wiki is correct. For the most part I think this is a safe assumption but genres could be misclassified, episodes could be mislabeled, etc.
3.  Each episode paired would score the same IMDb rating if voted on together. This is probably not true but IMDb only takes ratings for episodes paired in a 22-ish minute airing.

# Whatchyu doin'?

The rest of this report is organized by questions I thought were interesting or plots I thought might be cool.

```{r, include=FALSE}
library(tidyverse)
library(showtext)
library(lubridate)
library(ggtext)
```

```{r, include=FALSE}
songs <- read_csv("pnf_songs.csv")
```

```{r, include=FALSE}
small_plot_text <- element_text(family = "Graphite Std Bold", 
                                 size = 15,
                                 color = "black")

big_plot_text <- element_text(face = "bold", 
                                  family = "Ferbtastic",
                                  size = 30)
```

## What are the most common genres?

The following plot shows the top most common song genres in Phineas and Ferb and their prevalence among different seasons. Even though this is probably the most basic question to ask about the songs, I thought it was interesting because even Dan Povenmire himself might not have an idea of just how many pop and rock songs there are (or how few jazz songs). I also don't mean to imply that there needs to be an equal distribution of genres, I believe the song should fit the context of the episode around it.

```{r, echo=FALSE, warning=FALSE, fig.width=8, dpi=300}
songs %>% 
  mutate(genre = str_to_lower(genre)) %>% 
  separate(genre, into = c("g1", "g2", "g3", "g4", "g5"),
           sep = "[,/]") %>% 
  pivot_longer(g1:g5, "col", values_to = "genre") %>% 
  mutate(genre = str_trim(genre)) %>% 
  filter(!is.na(genre)) %>% 
  mutate(season = substr(pc, 1, 1)) %>% 
  group_by(genre) %>% 
  mutate(genre_count = n()) %>% 
  filter(genre_count > 10) %>% 
  count(genre, season, sort = TRUE) %>% 
  na.omit(season) %>% 
  group_by(genre) %>% 
  mutate(genre_count = sum(n)) %>% 
  ungroup() %>% 
  mutate(genre = fct_reorder(genre, genre_count)) %>% 
  mutate(season = case_when(season == "M" ~ "movie", 
                            TRUE ~ season)) %>% 
  mutate(season = fct_relevel(season, "movie", "4", "3", "2", "1")) %>% 
  ggplot(aes(y = genre, x = n, fill = season)) + 
  geom_col(color = "black") + 
  theme_minimal() + 
  scale_fill_brewer(palette = "YlOrRd", 
                    guide = guide_legend(reverse = TRUE)) + 
  labs(title = "Pop, Rock, and Show Tunes are the most\ncommon genre of P&F songs",
       caption = "\nData from phineasandferb.fandom.com  |  Scraped and Visualized by @PhilliPhresh") + 
  theme(axis.title = element_blank(), 
        panel.grid.major = element_blank(), 
        plot.title.position = "plot",
        plot.background = element_rect(fill = "#ffeac9"),
        plot.title = big_plot_text,
        axis.text = small_plot_text,
        legend.text = small_plot_text,
        legend.title = small_plot_text, 
        plot.caption = element_text(hjust = 1),
        plot.caption.position = "plot", 
        legend.direction = "horizontal", 
        legend.position = c(.65, .15))
```
